Let's talk about Thurstone \& Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2025-01-23 _tex/paper2.tex
Built site for gh-pages

Numerous studies have documented the effectiveness of CJ in assessing
traits and competencies over the past decade. These studies have
highlighted three aspects of the method's effectiveness: its
reliability, validity, and practical applicability. Research on
reliability suggests that CJ requires a relatively modest number of
pairwise comparisons \citep{Verhavert_et_al_2019, Crompvoets_et_al_2022}
to generate trait scores that are as precise and consistent as those
generated by other assessment methods
\citep{Coertjens_et_al_2017, Goossens_et_al_2018, Bouwer_et_al_2023}. In
addition, the evidence suggests that the reliability and time efficiency
of CJ are comparable, if not superior, to those of other assessment
methods when employing adaptive comparison algorithms
\citep{Pollitt_2012b, Verhavert_et_al_2022, Mikhailiuk_et_al_2021}.
Meanwhile, research on the validity of CJ scores indicates their
capacity to represent the traits under measurement accurately
\citep{Whitehouse_2012, vanDaal_et_al_2016, Lesterhuis_2018_thesis, Bartholomew_et_al_2018, Bouwer_et_al_2023}.
Moreover, research on CJ's practical applicability highlights its
versatility across both educational and non-educational contexts
\citep{Kimbell_2012, Jones_et_al_2015, Bartholomew_et_al_2018, Jones_et_al_2019, Marshall_et_al_2020, Bartholomew_et_al_2020, Boonen_et_al_2020}.

Nevertheless, despite the increasing number of CJ studies, the
prevalence of unsystematic and fragmented research approaches has left
several critical issues unaddressed. The present study primarily focuses
on two issues: the overreliance on Thurstone's Case V assumptions in the
statistical analysis of CJ data and the apparent disconnect between CJ's
approach to trait measurement and hypothesis testing. The following
sections begin with a brief overview of Thurstone's theory followed by a
detailed examination of these issues. Subsequently, the study introduces
a theoretical model for CJ that builds upon Thurstone's theory,
alongside its statistical translation, designed to address the two
concerns simultaneously.

@@

Figure~\ref{fig-discriminal_process} illustrates hypothetical
discriminal processes along a quality trait continuum for two written
texts. The figure indicates that the modal discriminal process for Text
B is positioned further along the continuum than that of Text A
\((T_{B} > T_{A})\), suggesting that Text B exhibits higher quality.
Additionally, the figure highlights that Text B has a broader
distribution compared to Text A, which arises from its larger
discriminal dispersion \((\sigma_{B} > \sigma_{A})\).

@@

However, given that the individual discriminal processes of the stimuli
are not directly observable, the theory introduces the \emph{law of
comparative judgment}. This law posits that in pairwise comparisons, a
judge perceives the stimulus with a discriminal process positioned
further along the trait continuum as possessing more of the trait
\citep[pp.~251]{Bramley_2008}. This suggests that the relative distance
between stimuli, rather than their absolute positions on the continuum,
likely defines the outcome of pairwise comparisons. Indeed, the theory
assumes that the difference between the underlying discriminal processes
of the stimuli, referred to as the \emph{discriminal difference},
determines the observed dichotomous outcome. Furthermore, the theory
assumes that because the individual discriminal processes form a Normal
distribution on the continuum, the discriminal difference will also
conform to a Normal distribution \citep{Andrich_1978}. In this
distribution, the mode (mean) represents the relative separation between
the stimuli, and its dispersion indicates the variability of that
separation.

@@

This section examines the two critical issues in the CJ literature that
serve as the primary focus of the present study. The first is related to
the overreliance on Thurstone's Case V assumptions in the statistical
analysis of CJ data. The second concern with the apparent disconnect
between CJ's approach to trait measurement and hypothesis testing.

@@

According to the theory, discrepancies in the discriminal dispersions of
stimuli shape the distribution of the discriminal difference, exerting a
direct influence on the outcome of pairwise comparisons.
Figure~\ref{fig-dispersion} presents a thought experiment to illustrate
this idea. In this experiment, a researcher can observe the discriminal
processes for the texts depicted in
Figure~\ref{fig-discriminal_process}. Furthermore, the figure assumes
that the discriminal dispersion for Text A remains constant and that the
texts are uncorrelated \((\rho=0)\). The figure reveals that an increase
in the uncertainty associated with the perception of Text B in
comparison to Text A, \((\sigma_{B}-\sigma_{A})\), broadens the
distribution of their discriminal difference. This broadening affects
the probability area where the discriminal difference distinctly favors
Text B over Text A, expressed as \(P(B > A)\), ultimately influencing
the comparison outcome. Additionally, the figure reveals that when the
discriminal dispersions of the texts are equal
\((\sigma_{B}-\sigma_{A}=0)\), the discriminal difference is more likely
to favor Text B over Text A (shaded gray area), compared to situations
where their dispersions differ.

@@

In experimental practice, however, this process occurs in reverse.
Researchers first observe the comparison outcome and then use the BTL
model to infer the discriminal difference between the stimuli and their
respective discriminal processes \citep[pp.~373]{Thurstone_1927a}.
Therefore, the outcome's ability to reflect the ``true'' differences
between stimuli largely depends on the validity of the model's
assumptions \citep[pp.~150]{Kohler_et_al_2019}, particularly the
assumption of equal dispersions. For instance, when researchers observe
a sample of outcomes favoring Text B over Text A and correctly assume
equal dispersions between the texts, the BTL model estimates a
discriminal difference distribution that accurately represents the
``true'' discriminal difference of the texts. This scenario is
illustrated in Figure~\ref{fig-dispersion}, where the model's
discriminal difference distribution aligns with the ``true''
distribution, represented by the thick continuous line corresponding to
\(\sigma_{B}-\sigma_{A}=0\). The accuracy of these discriminal
difference ensures reliable estimates for the texts' discriminal
processes {(citation needed?)}.

@@

Thus, model misspecification, in the form of an erroneous assumption of
equal dispersions between stimuli, can give rise to significant
statistical and measurement issues. For instance, the model may
overestimate the degree to which the outcome accurately reflects the
``true'' discriminal differences between stimuli. This overestimation
can result in researchers drawing spurious conclusions about these
differences \citep[pp.~370]{McElreath_2020} and, by extension, about the
underlying discriminal processes of stimuli. Figure~\ref{fig-dispersion}
also illustrates this issue when the model's discriminal difference
distribution aligns with the thick continuous line for
\(\sigma_{B}-\sigma_{A}=0\), while the ``true'' discriminal difference
follows any discontinuous line where \(\sigma_{B}-\sigma_{A} \neq 0\).
Additionally, if researchers recognize that misfit statistics highlight
these critical differences in dispersions, the conventional CJ practice
of excluding stimuli based on these statistics
\citep{Pollitt_2012a, Pollitt_2012b, vanDaal_et_al_2016, Goossens_et_al_2018}
can unintentionally discard valuable information. Such exclusions can
introduce bias into trait estimates
\citep[chap.~12]{Zimmerman_1994, McElreath_2020}. The direction and
magnitude of these biases are often unpredictable, as they depend on
which stimuli are excluded from the analysis.

@@

The correlation, represented by the symbol \(\rho\), measures how much a
judge's perception of a specific trait in one stimulus depends on their
perception of the same trait in another. As with the discriminal
dispersions, this correlation shapes the distribution of the discriminal
difference, directly impacting the outcomes of pairwise comparisons.
Figure~\ref{fig-correlation} presents a similar thought experiment as in
Section~\ref{sec-theory-issue1a} to illustrate this idea. The
illustration now assumes that the discriminal dispersions for both texts
remain constant. The figure reveals that as the correlation between the
texts increases, the distribution of their discriminal difference
becomes narrower. This narrowing affects the area under the curve where
the discriminal difference distinctly favors Text B over Text A, denoted
as \(P(B > A)\), thus influencing the comparison outcome. Furthermore,
the figure shows that when two texts are independent or uncorrelated
\((\rho=0)\), their discriminal difference is less likely to favor Text
B over Text A (shaded gray area) compared to scenarios where the texts
are highly correlated.

@@

Notably, Thurstone attributed the independence of stimuli to the
cancellation of potential judges' biases. He argued that this
cancellation resulted from two opposing and equally weighted effects
occurring during pairwise comparisons \citep[pp.~268]{Thurstone_1927b}.
\citet{Andrich_1978} provided a mathematical demonstration of this
cancellation using the BTL model under the assumption of discriminal
processes with additive biases. However, it is easy to imagine at least
two scenarios in which the zero correlation assumption is almost
certainly invalid: when the pairwise comparison involves
multidimensional, complex traits with heterogeneous stimuli and when an
additional hierarchical structure is relevant to the stimuli.

@@

In the second scenario, the shared context or inherent connections
created by additional hierarchical structures may further introduce
dependencies between stimuli, a statistical phenomenon commonly known as
clustering \citep{Everitt_et_al_2010}. Despite the CJ literature
acknowledging the existence of such hierarchical structures, the
statistical handling of this additional source of dependence between
stimuli has been inadequate. For instance, when CJ data incorporates
multiple samples of stimuli from the same individuals, researchers
frequently rely on (average) estimated BTL scores to conduct subsequent
analyses and tests at the individual hierarchical level
\citep{Bramley_et_al_2019, Boonen_et_al_2020, Bouwer_et_al_2023, vanDaal_et_al_2017, Jones_et_al_2019, Gijsen_et_al_2021}.
However, this approach can introduce additional statistical and
measurement issues, which we discuss in greater detail in
Section~\ref{sec-theory-issue2}.

@@

The misspecification may arise from neglecting additional relevant
traits, excluding judges based on misfit statistics, or ignoring
hierarchical (grouping) structures. Neglecting relevant traits, such as
judges' biases, can cause dimensional mismatches in the BTL model,
artificially inflating the trait's reliability
\citep[pp.~341]{Hoyle_et_al_2023} or, worse, introducing bias into the
trait's estimates \citep{Ackerman_1989}. Excluding judges based on
misfit statistics risks discarding valuable information, which may
further bias the trait's estimates
\citep[chap.~12]{Zimmerman_1994, McElreath_2020}. Finally, ignoring
hierarchical structures may reduce the precision of model parameter
estimates, potentially amplifying the overestimation of the trait's
reliability \citep[pp.~482]{Hoyle_et_al_2023}.

@@

In aggregate, researchers' inadequate handling of violations to the
assumptions of equal dispersion and zero correlation between stimuli,
coupled with the apparent disconnect between CJ's approach to trait
measurement and hypothesis testing, can potentially compromise the
reliability of the trait estimates and, by extension, their validity
\citep[pp.~2]{Perron_et_al_2015}. Consequently, adopting a more
systematic and integrated approach to handling these assumptions and
examining the factors influencing CJ experiments could offer several
statistical and measurement benefits, including the ability to address
these issues.

@@

This section uses the structural approach to causal inference
\citep{Pearl_2009, Pearl_et_al_2016} to articulate a theoretical model
that captures the core principles of Thurstone's theory. The model also
incorporates various assessment design features relevant to CJ
experiments, such as the selection of judges, stimuli, and comparisons.
Finally, the section employs Bayesian inference methods to transform
these theoretical and practical elements into a statistical model that
facilitates the analysis of pairwise comparison data. See
Section~\ref{sec-appendix} for an overview of the statistical and causal
inference concepts required for the development of this section.

@@

The theoretical model uses structural causal models (SCMs) and directed
acyclic graphs (DAGs)
\citep{Pearl_2009, Pearl_et_al_2016, Gross_et_al_2018, Neal_2020} to
formally and graphically represent the assumed causal structure of the
CJ system. First, \emph{the population model} is created to represent a
conceptual population of CJ experiments. The model then integrates
various assessment design features relevant to CJ experiments, leading
to the development of \emph{the sample-comparison model}.

@@

In this section we focus on the identification of several parameters of
interest from the theoretical model, together with various assessment
design features relevant to CJ experiments. The main reason to focus on
identification is that if these parameters cannot be identified using
``infinite'' and error free data, we certainly cannot learn anything
about it using finite data. So, for the time being, we assume large
samples and error-free measurements \citep[pp.~5]{Schuessler_et_al_2023}

@@

\section{Discussion}\label{sec-discuss}

\subsection{Findings}\label{sec-discuss-finding}

\subsection{Limitations and further
research}\label{sec-discuss-limitations}

>>>>

Numerous studies have documented the effectiveness of CJ in assessing
traits and competencies over the past decade. These studies have
highlighted three aspects of the method's effectiveness: its
reliability, validity, and practical applicability. Research on
reliability suggests that CJ requires a relatively modest number of
pairwise comparisons \citep{Verhavert_et_al_2019, Crompvoets_et_al_2022}
to generate trait scores that are as precise and consistent as those
generated by other assessment methods
\citep{Coertjens_et_al_2017, Goossens_et_al_2018, Bouwer_et_al_2023}. In
addition, the evidence suggests that the reliability and time efficiency
of CJ are comparable, if not superior, to those of other assessment
methods when employing adaptive comparison algorithms
\citep{Pollitt_2012b, Verhavert_et_al_2022, Mikhailiuk_et_al_2021}.
Meanwhile, research on the validity of CJ scores indicates their
capacity to represent accurately the traits under measurement
\citep{Whitehouse_2012, vanDaal_et_al_2016, Lesterhuis_2018_thesis, Bartholomew_et_al_2018, Bouwer_et_al_2023}.
Moreover, research on CJ's practical applicability highlights its
versatility across both educational and non-educational contexts
\citep{Kimbell_2012, Jones_et_al_2015, Bartholomew_et_al_2018, Jones_et_al_2019, Marshall_et_al_2020, Bartholomew_et_al_2020, Boonen_et_al_2020}.

Nevertheless, despite the growing number of CJ studies, unsystematic and
fragmented research approaches have left several critical issues
unresolved. This study examines two such problems that can compromise
the reliability and validity of CJ's trait estimates. The first issue
concerns the overreliance on Thurstone's Case V assumptions in the
statistical analysis of CJ data. The second involves the apparent
disconnect between CJ's approach to trait measurement and hypothesis
testing.

The study has five main sections. It begins with a brief overview of
Thurstone's theory in Section~\ref{sec-thurstone_theory}, followed by a
detailed examination of these issues in Section~\ref{sec-theory-issues}.
In Section~\ref{sec-theory}, the study introduces a theoretical model
for CJ that builds upon Thurstone's theory, along with its statistical
translation, designed to address both issues simultaneously. Finally,
Section~\ref{sec-discussion} discusses the concepts developed in the
study, their implications, and avenues for future exploration, while
Section~\ref{sec-conclusion} presents the conclusions.

@@

Figure~\ref{fig-discriminal_process} illustrates the hypothetical
discriminal processes along a quality trait continuum for two written
texts. The figure indicates that the modal discriminal process for Text
B is positioned further along the continuum than that of Text A
\((T_{B} > T_{A})\), suggesting that Text B exhibits higher quality.
Additionally, the figure highlights that Text B has a broader
distribution compared to Text A, which arises from its larger
discriminal dispersion \((\sigma_{B} > \sigma_{A})\).

@@

However, since the individual discriminal processes of the stimuli are
not directly observable, the theory introduces the \emph{law of
comparative judgment}. This law posits that in pairwise comparisons, a
judge perceives the stimulus with a discriminal process positioned
further along the trait continuum as possessing more of the trait
\citep[pp.~251]{Bramley_2008}. This suggests that the relative distance
between stimuli, rather than their absolute positions on the continuum,
likely defines the outcome of pairwise comparisons. Indeed, the theory
assumes that the difference between the underlying discriminal processes
of the stimuli, referred to as the \emph{discriminal difference},
determines the observed dichotomous outcome. Furthermore, the theory
assumes that because the individual discriminal processes form a Normal
distribution on the continuum, the discriminal difference will also
conform to a Normal distribution \citep{Andrich_1978}. In this
distribution, the mode (mean) represents the relative separation between
the stimuli, and its dispersion indicates the variability of that
separation.

@@

This section examines the two critical issues in CJ's theory that can
compromise the reliability and validity of CJ's trait estimates. In
Section~\ref{sec-theory-issue1}, the study examines the first issue: the
overreliance on Thurstone's Case V assumptions in the statistical
analysis of CJ data. In Section~\ref{sec-theory-issue2}, the study
explores the second issue: the apparent disconnect between CJ's approach
to trait measurement and hypothesis testing.

@@

According to the theory, discrepancies in the discriminal dispersions of
stimuli shape the distribution of the discriminal difference, exerting a
direct influence on the outcome of pairwise comparisons. A thought
experiment can illustrate this idea. In this experiment, the researcher
can observe the discriminal processes for the texts depicted in
Figure~\ref{fig-discriminal_process}. Furthermore, the experiment
assumes that the discriminal dispersion for Text A remains constant and
that the texts are uncorrelated \((\rho=0)\). In this scenario,
Figure~\ref{fig-dispersion} reveals that an increase in the uncertainty
associated with the perception of Text B in comparison to Text A,
\((\sigma_{B}-\sigma_{A})\), broadens the distribution of their
discriminal difference. This broadening affects the probability area
where the discriminal difference distinctly favors Text B over Text A,
expressed as \(P(B > A)\), ultimately influencing the comparison
outcome. Additionally, the figure reveals that when the discriminal
dispersions of the texts are equal \((\sigma_{B}-\sigma_{A}=0)\), the
discriminal difference is more likely to favor Text B over Text A
(shaded gray area), compared to situations where their dispersions
differ.

@@

In experimental practice, however, the thought experiment occurs in
reverse. Researchers first observe the comparison outcome and then use
the BTL model to infer the discriminal difference between the stimuli
and their respective discriminal processes
\citep[pp.~373]{Thurstone_1927a}. Therefore, the outcome's ability to
reflect the ``true'' differences between stimuli largely depends on the
validity of the model's assumptions \citep[pp.~150]{Kohler_et_al_2019},
particularly the assumption of equal dispersions. For instance, when
researchers observe a sample of outcomes favoring Text B over Text A and
correctly assume equal dispersions between the texts, the BTL model
estimates a discriminal difference distribution that accurately
represents the ``true'' discriminal difference of the texts. This
scenario is illustrated in Figure~\ref{fig-dispersion}, when the model's
discriminal difference distribution aligns with the ``true''
distribution, represented by the thick continuous line corresponding to
\(\sigma_{B}-\sigma_{A}=0\). The accuracy of the discriminal difference
then ensures reliable estimates for the texts' discriminal processes
{(citation needed?)}.

@@

Thus, model misspecification, in the form of an erroneous assumption of
equal dispersions between stimuli, can give rise to significant
statistical and measurement issues. For instance, the model may
overestimate the degree to which the outcome accurately reflects the
``true'' discriminal differences between stimuli. This overestimation
can result in researchers drawing spurious conclusions about these
differences \citep[pp.~370]{McElreath_2020} and, by extension, about the
underlying discriminal processes of stimuli. Figure~\ref{fig-dispersion}
also illustrates this issue when the model's discriminal difference
distribution aligns with the thick continuous line for
\(\sigma_{B}-\sigma_{A}=0\), while the ``true'' discriminal difference
follows any discontinuous line where \(\sigma_{B}-\sigma_{A} \neq 0\).
Additionally, if researchers recognize that misfit statistics highlight
these critical differences in dispersions, the conventional CJ practice
of excluding stimuli based on these statistics
\citep{Pollitt_2012a, Pollitt_2012b, vanDaal_et_al_2016, Goossens_et_al_2018}
can unintentionally discard valuable information, introducing bias into
trait estimates \citep[chap.~12]{Zimmerman_1994, McElreath_2020}. The
direction and magnitude of these biases are often unpredictable, as they
depend on which stimuli are excluded from the analysis.

@@

The correlation, represented by the symbol \(\rho\), measures how much a
judge's perception of a specific trait in one stimulus depends on their
perception of the same trait in another. As with the discriminal
dispersions, this correlation shapes the distribution of the discriminal
difference, directly impacting the outcomes of pairwise comparisons. A
similar thought experiment, as the one in
Section~\ref{sec-theory-issue1a}, can illustrate this idea. The
experiment now assumes that the discriminal dispersions for both texts
remain constant. Figure~\ref{fig-correlation} reveals that as the
correlation between the texts increases, the distribution of their
discriminal difference becomes narrower. This narrowing affects the area
under the curve where the discriminal difference distinctly favors Text
B over Text A, denoted as \(P(B > A)\), thus influencing the comparison
outcome. Furthermore, the figure shows that when two texts are
independent or uncorrelated \((\rho=0)\), their discriminal difference
is less likely to favor Text B over Text A (shaded gray area) compared
to scenarios where the texts are highly correlated.

@@

Notably, Thurstone attributed the lack of correlation (independence)
between stimuli to the cancellation of potential judges' biases. He
argued that this cancellation resulted from two opposing and equally
weighted effects occurring during pairwise comparisons
\citep[pp.~268]{Thurstone_1927b}. \citet{Andrich_1978} provided a
mathematical demonstration of this cancellation using the BTL model
under the assumption of discriminal processes with additive biases.
However, it is easy to imagine at least two scenarios in which the zero
correlation assumption is almost certainly invalid: when the pairwise
comparison involves multidimensional, complex traits with heterogeneous
stimuli and when an additional hierarchical structure is relevant to the
stimuli.

@@

In the second scenario, the shared context or inherent connections
created by additional hierarchical structures may further introduce
dependencies between stimuli, creating a statistical phenomenon known as
clustering \citep{Everitt_et_al_2010}. Although the CJ literature
acknowledges the existence of such hierarchical structures, the
statistical handling of this additional source of dependence between
stimuli has been inadequate. For instance, when CJ data incorporates
multiple samples of stimuli from the same individuals, researchers
frequently rely on (average) estimated BTL scores to conduct subsequent
analyses and tests at the individual hierarchical level
\citep{Bramley_et_al_2019, Boonen_et_al_2020, Bouwer_et_al_2023, vanDaal_et_al_2017, Jones_et_al_2019, Gijsen_et_al_2021}.
This approach, however, can introduce additional statistical and
measurement issues, which we discuss in greater detail in
Section~\ref{sec-theory-issue2}.

@@

The model misspecification may arise from neglecting additional relevant
traits, excluding judges based on misfit statistics, or ignoring
hierarchical (grouping) structures. Neglecting relevant traits, such as
judges' biases, can cause dimensional mismatches in the BTL model,
artificially inflating the trait's reliability
\citep[pp.~341]{Hoyle_et_al_2023} or, worse, introducing bias into the
trait's estimates \citep{Ackerman_1989}. Excluding judges based on
misfit statistics risks discarding valuable information, which may
further bias the trait's estimates
\citep[chap.~12]{Zimmerman_1994, McElreath_2020}. Finally, ignoring
hierarchical structures may reduce the precision of model parameter
estimates, potentially overestimating the trait's reliability
\citep[pp.~482]{Hoyle_et_al_2023}.

@@

In aggregate, researchers' inadequate handling of violations to the
assumptions of equal dispersion and zero correlation between stimuli,
coupled with the apparent disconnect between CJ's approach to trait
measurement and hypothesis testing, can potentially compromise the
reliability of the trait estimates and, by extension, their validity
\citep[pp.~2]{Perron_et_al_2015}. Consequently, adopting a more
systematic and integrated approach to handling these assumptions and
examining the factors influencing CJ experiments could offer several
statistical and measurement benefits, including the ability to address
the aforementioned issues.

@@

The objective of this section is twofold. First, it aims to refine CJ's
theoretical model using the \emph{structural approach} to causal
inference. This approach, as outlined in
Section~\ref{sec-theory-theoretical}, defines a model that integrates
the core principles of Thurstone's theory and key assessment design
features relevant to CJ experiments. Second,
Section~\ref{sec-theory-statistics} uses \emph{Bayesian inference
methods} to convert these theoretical and practical elements into a
statistical model that supports the analysis of paired comparison data.

@@

The \emph{structural approach} to causal inference provides a formal
framework to identify causes and estimate their effects using data. The
approach uses structural causal models (SCMs) and directed acyclic
graphs (DAGs)
\citep{Pearl_2009, Pearl_et_al_2016, Gross_et_al_2018, Neal_2020} to
formally and graphically represent the assumed causal structure of a
system. Essentially, SCMs and DAGs serve as \emph{conceptual
(theoretical) models} that facilitate valid causal inference. They
achieve this by helping researchers to determine which statistical
models can identify an estimand, assuming the depicted causal structure
is correct \citep{McElreath_2020}. \emph{Identification} determines
whether an estimator can accurately compute an estimand based on its
assumptions, regardless of random variability
\citep[pp.~4]{Schuessler_et_al_2023}.

SCMs are formal mathematical models defined by a set of
\emph{endogenous} variables, a set of \emph{exogenous} variables, and a
set of functions \citep{Pearl_2009, Cinelli_et_al_2020}. Endogenous
variables are those whose causal mechanisms a researcher chooses to
model \citep{Neal_2020}. In contrast, exogenous variables represent
\emph{errors} or \emph{disturbances} arising from omitted factors that
the investigator chooses not to model explicitly
\citep[pp.~27,68]{Pearl_2009}. Lastly, the functions, referred to as
\emph{structural equations}, express the endogenous variables as
non-parametric functions of other variables. These functions use the
symbol `\(:=\)' to denote the asymmetrical causal dependence between
variables and the symbol `\(\:\bot\:\)' to represent
\emph{d-separation}, a concept akin to (conditional) independence.

Notably, every SCM has an associated DAG
\citep{Pearl_et_al_2016, Cinelli_et_al_2020}. A DAG is a graph
consisting of nodes connected by edges, where the nodes represent random
variables. The term \emph{directed} means that the edges extend from one
node to another, with arrows indicating the direction of causal
influence. The term \emph{acyclic} implies that the causal influences do
not form loops, ensuring the influences do not cycle back on themselves
\citep{McElreath_2020}. DAGs represent observed variables as solid black
circles, while they use open circles for unobserved (latent) variables
\citep{Morgan_et_al_2014}. While the \emph{standard representation} of
DAGs typically omits exogenous variables for simplicity, the
\emph{magnified representation} includes them, offering offering the
advantage of highlighting potential issues related to conditioning and
confounding \citep{Cinelli_et_al_2020}.

Therefore, this section leverages SCMs and DAGs to represent the
conceptual population of CJ experiments, referred to as \emph{the
population model} in Section~\ref{sec-theory-theoretical_P}, and later
integrates various assessment design features relevant to CJ
experiments. This process leads to the development of \emph{the
sample-comparison model} in Section~\ref{sec-theory-theoretical_SC}.

@@

@@

\section{Discussion}\label{sec-discussion}

\subsection{Findings}\label{sec-discussion-finding}

\subsection{Limitations and further
research}\label{sec-discussion-limitations}
